---
title: "Coporate_PAC"
author: "Lim"
date: "March 11, 2020"
output: rmarkdown::github_document
header-includes:
  - \usepackage{color}
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract
Do the ideological biases of corporate boards constrain the giving of corporate PACs? Previous literature shows that corporate PACs' board of directors are ideologically heterogeneous and PACs are strategically centrists in terms of their profile of campaign contribution. However, there is little literature on why and how corporate PACs' political spending can be constrained by corporate elites. My finding shows that the more ideologically homogeneous boardroom, the more corporate PAC spending. This finding can be explained by a theory of internal constraints in interest groups' activities in that donors of interest groups can constrain the groups' political contributions. This paper's finding suggests that corporate PACs are not only constrained by the donors but also they can be constrained by corporate elites in that the board of directors can influence PACs' amount of spending. This paper's findings can help us to better understand why there is so little money in politics and why corporate PACs are concentrated in the center of ideological spectrum.


```{r, message=FALSE,warning=FALSE,echo=F}

library(ggplot2)
library(dplyr)
library(MASS)
library(pscl)

dat<-read.csv("C:/Users/wooki/Dropbox/APB/final_data2.csv",header=TRUE)


##########################################
#summary data
dat <- within(dat, {
  corp.name <- factor(corp.name)
})
```

```{r, message=FALSE,warning=FALSE,echo=F}
#axis dollar tick
fancy_scientific <- function(l) {
  # turn in to character string in scientific notation
  l <- format(l, scientific = TRUE)
  # quote the part before the exponent to keep all the digits
  l <- gsub("^(.*)e", "'\\1'e", l)
  # turn the 'e+' into plotmath format
  l <- gsub("e", "%*%10^", l)
  # return this as an expression
  parse(text=l)
}

####################
human_numbers <- function(x = NULL, smbl ="", signif = 1){
  humanity <- function(y){
    
    if (!is.na(y)){
      tn <- round(abs(y) / 1e12, signif)
      b <- round(abs(y) / 1e9, signif)
      m <- round(abs(y) / 1e6, signif)
      k <- round(abs(y) / 1e3, signif)
      
      if ( y >= 0 ){
        y_is_positive <- ""
      } else {
        y_is_positive <- "-"
      }
      
      if ( k < 1 ) {
        paste0( y_is_positive, smbl, round(abs(y), signif ))
      } else if ( m < 1){
        paste0 (y_is_positive, smbl,  k , "k")
      } else if (b < 1){
        paste0 (y_is_positive, smbl, m ,"m")
      }else if(tn < 1){
        paste0 (y_is_positive, smbl, b ,"bn")
      } else {
        paste0 (y_is_positive, smbl,  comma(tn), "tn")
      }
    } else if (is.na(y) | is.null(y)){
      "-"
    }
  }
  
  sapply(x,humanity)
}

human_gbp   <- function(x){human_numbers(x, smbl = "£")}
human_usd   <- function(x){human_numbers(x, smbl = "$")}
human_euro  <- function(x){human_numbers(x, smbl = "???")} 
human_num   <- function(x){human_numbers(x, smbl = "")} 

```

## Including Plots

You can also embed plots, for example:

```{r, message=FALSE,warning=FALSE,echo=F}

p1<-ggplot(dat, aes(sum12, fill = sector)) + geom_density() + 
  facet_grid(sector ~ ., margins = TRUE, scales = "free")+
  scale_x_continuous(labels = human_usd)+
  scale_y_continuous(labels = fancy_scientific)+
  #ggtitle("Distribution of croporates PAC's spending by sector")+
  labs(x="Corporate PAC's total spending",y="Density")+
  theme_classic()
suppressMessages(p1+theme(plot.title=element_text(hjust=0.5)))
```


```{r , message=FALSE,warning=FALSE,echo=F}

p3<-ggplot(dat, aes(var.dime.cfscore)) + geom_density()+
  annotate("text", x = 0.147641, y = 1, angle = 90, label = paste("Democrat"), 
           vjust = 1.2, parse = TRUE)+
  annotate("text", x = 0.05295302, y = 1.3, angle = 90, label = paste("Republican"), vjust = 1.2)+
  annotate("text", x = 0.050499315, y = 0.8, angle = 90, label = "Sierra Club", vjust = 1.2)+
  annotate("text", x = 0.954, y = 1, angle = 90, label = "Congress", vjust = 1.2)+
  annotate("text", x = 0.41, y = 1, angle = 90, label = paste("Google"), 
           vjust = 1.2, parse = TRUE)+
  labs(x="Variance of CFscores within a group",y="Density")+
  #ggtitle("Distribution of variance BOD CFscores within corporates")+
  theme_classic()
suppressMessages(p3)
```


```{r , message=FALSE,warning=FALSE,echo=F}
bod<-read.csv("C:/Users/wooki/Dropbox/APB/bod.csv",header=TRUE)
data<-subset(bod, bod$corp.name=="google")
(ids <- sample(unique(data$corp.name), 1))

#narrow your data set
data2 <- data[data$corp.name %in% ids, ]
#write.csv(data2, "group compare.csv")
data2<-read.csv("C:/Users/wooki/Dropbox/APB/group compare.csv",header=TRUE)
my_mean=aggregate(data2$dime.cfscore , by=list(data2$corp.name) , mean, na.rm=T) ; colnames(my_mean)=c("names" , "mean")
my_sd=aggregate(data2$dime.cfscore , by=list(data2$corp.name) , sd, na.rm=T) ; colnames(my_sd)=c("names" , "sd")
my_info=merge(my_mean , my_sd , by.x=1 , by.y=1)


# Make the plot
g<-ggplot(data2) + 
  geom_point(aes(y = corp.name, x = dime.cfscore) ,size = 1) + 
  geom_point(data = my_info, aes(y=names , x = mean) , colour = "red" , size = 1) +
  labs(x="CFscores",y="")+
  theme_classic()

suppressMessages(g)
```



```{r , message=FALSE,warning=FALSE,echo=F}
p4<-ggplot(dat, aes(y=sum12, x = var.dime.cfscore)) + geom_point() +
  labs(x="Variance of BOD CFscores",y="Corporate PAC's total spending")+
  scale_y_continuous(labels = human_usd)+
  #ggtitle("Corporate PAC's total spending and Varince of BOD's CFscores")+
  theme_classic()
suppressMessages(p4+theme(plot.title=element_text(hjust=0.5)))

```

```{r, message=FALSE,warning=FALSE,echo=T}
m1 <- glm.nb(sum12 ~ var.dime.cfscore+sector+mean.dime.cfscore+p.female, data = dat)
summary(m1)
```
